---
title: "A3"
output: html_document
date: "2025-05-15"
---



```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(summarytools)
library(DataExplorer)
library(corrplot)
library(dplyr)
library(zoo)  #rollmean smoothness
library(MLmetrics)
library(caret)
library(randomForest)
library(pROC)
library(ROSE)
library(ggrepel)
library(viridis)

stroke_data <- read.csv("D:/Desktop/2025S1/HTIN5006/A3/brain_stroke.csv")
stroke_data
```

# EDA
```{r cars}
#  check data structure
str(stroke_data)

# 
head(stroke_data)

# data dimension
dim(stroke_data)

#  missing values
colSums(is.na(stroke_data))

# Descriptive statistics (numerical variables)
summary(select_if(stroke_data, is.numeric))

# Descriptive statistics (categorical variables)
lapply(select_if(stroke_data, is.factor), table)
```

```{r}
# Calculate the number of missing values in each column
missing_counts <- colSums(is.na(stroke_data))
print(missing_counts)

# Calculate the proportion of missing values in each column
missing_ratios <- colMeans(is.na(stroke_data))
print(missing_ratios)

# Summarize and display the missing situations (only show the missing columns)
missing_summary <- data.frame(
  Variable = names(missing_counts),
  MissingCount = missing_counts,
  MissingRatio = missing_ratios
)

missing_summary[missing_summary$MissingCount > 0, ]
```


```{r}
# Convert the categorical variable to factor
stroke_data$ever_married <- as.factor(stroke_data$ever_married)
stroke_data$work_type <- as.factor(stroke_data$work_type)
stroke_data$Residence_type <- as.factor(stroke_data$Residence_type)
stroke_data$smoking_status <- as.factor(stroke_data$smoking_status)
stroke_data$stroke <- as.factor(stroke_data$stroke)
```


```{r}
#Some visual analyses

# Distribution of numerical variables
plot_histogram(stroke_data)

# Frequency graph of category variables
plot_bar(stroke_data)

# Correlation heat map (numerical variable)
numeric_data <- select_if(stroke_data, is.numeric)
corr <- cor(na.omit(numeric_data))
corrplot(corr, method = "color", addCoef.col = "black")


```

```{r}
#Examine the proportion of strokes in each category within each categorical variable
cat_vars <- c("gender", "ever_married", "work_type", "Residence_type", "smoking_status")

for (var in cat_vars) {                     # Calculate the proportion of stroke under each categorical variable in a loop
  print(paste("stroke proportion", var, "group："))
  
    stroke_prop <- stroke_data %>%
      group_by(!!sym(var), stroke) %>%     # Grouped by variables and stroke status
      summarise(count = n()) %>%          # Calculate frequency
      mutate(prop = count / sum(count)) %>%  # Calculate the proportion
      filter(stroke == 1) %>%              # Only select the group =1
      select(!!sym(var), prop)            # Only retain the variable names and proportions
    
  print(stroke_prop)
}

```

```{r}
#The relationship between stroke and some factors
# stroke & age 
ggplot(stroke_data, aes(x = age, fill = stroke)) +
  geom_histogram(binwidth = 5, position = "dodge") +
  labs(title = "age distribution vs stroke occurs", x = "age", y = "people") +
  theme_minimal()

# stroke & gender
ggplot(stroke_data, aes(x = gender, fill = stroke)) +
  geom_bar(position = "fill") +
  labs(title = "gender vs stroke", y = "proportion") +
  theme_minimal()

# stroke & smoking_status 
ggplot(stroke_data, aes(x = smoking_status, fill = stroke)) +
  geom_bar(position = "fill") +
  labs(title = "smoking status vs stroke", y = "proportion") +
  theme_minimal()
```




#Exploration of the Relationship between stroke Probability and Age

```{r}
# Convert all categorical variables in the training data into factors (especially gender)
factor_vars <- c("gender", "hypertension", "heart_disease", "ever_married",
                 "work_type", "Residence_type", "smoking_status")

for (var in factor_vars) {
  stroke_data[[var]] <- factor(stroke_data[[var]])
}
train_index <- createDataPartition(stroke_data$stroke, p = 0.7, list = FALSE)
train_data <- stroke_data[train_index, ]
test_data <- stroke_data[-train_index, ]

# Generate balanced training data
set.seed(123)
balanced_train_data <- ROSE(stroke ~ ., data = train_data, seed = 1)$data

# Check the category distribution of the new dataset
table(balanced_train_data$stroke)
```
```{r}
# Step 1: Extract variables
pca_data <- balanced_train_data[, c("age", "avg_glucose_level", "bmi")]

# Step 2: standardization
pca_scaled <- scale(pca_data)

# Step 3: PCA
pca_result <- prcomp(pca_scaled, center = TRUE, scale. = TRUE)

# Step 4: View the principal component variance explanation
summary(pca_result)

# Step 5: Check which variables each principal component is composed of (weights)
pca_result$rotation

# Step 6: Combine the principal component score with the stroke label to form a data frame
pca_df <- as.data.frame(pca_result$x)
pca_df$stroke <- as.factor(balanced_train_data$stroke)  # Make sure it is of the factor type

# Step 7: Visualize the principal components (PC1 vs PC2) and distinguish whether it is a stroke by color
library(ggplot2)

ggplot(pca_df, aes(x = PC1, y = PC2, color = stroke)) +
  geom_point(alpha = 0.7, size = 2) +
  labs(title = "PCA: PC1 vs PC2（age、glucose、bmi）",
       x = "PC1",
       y = "PC2",
       color = "Stroke") +
  theme_minimal()

```

```{r}
# Group the ages

stroke_data <- stroke_data %>%
  mutate(age_group = cut(age,
                         breaks = seq(0, 100, by = 10),
                         right = FALSE,
                         labels = paste(seq(0, 90, by = 10), seq(10, 100, by = 10)-1, sep = "-")))

# Calculate the probability of each group of strokes
age_stroke_rate <- stroke_data %>%
  group_by(age_group) %>%
  summarise(
    stroke_rate = mean(as.numeric(as.character(stroke))),
    count = n()
  )

# visualization
ggplot(age_stroke_rate, aes(x = age_group, y = stroke_rate)) +
  geom_col(fill = "tomato") +
  labs(title = "The incidence of stroke in different age groups", x = "age group", y = "probability of stroke") +
  theme_minimal()

```

#Add other variables to explore the interaction
```{r}
#The logistic regression model contains interaction terms: age × hypertension, age × smoking status
# Simplify the data (handle omissions and select the main variables)
stroke_data <- na.omit(stroke_data)

# Transform categorical variables
stroke_data$hypertension <- as.factor(stroke_data$hypertension)
stroke_data$heart_disease <- as.factor(stroke_data$heart_disease)
stroke_data$smoking_status <- as.factor(stroke_data$smoking_status)
stroke_data$stroke <- as.factor(stroke_data$stroke)

# Logistic regression model with interaction terms
model <- glm(stroke ~ age * hypertension + age * smoking_status + age * heart_disease,
             data = stroke_data, family = binomial)

summary(model)

```



#Random Forest analysis
```{r}
set.seed(123)


# Train the random forest model
Randomforest_model0 <- randomForest(
  stroke ~ age + hypertension + heart_disease + ever_married + work_type +
    Residence_type + avg_glucose_level + bmi + smoking_status + gender,
  data = balanced_train_data,
  ntree = 500,         # The number of trees
  mtry = 3,            # The number of features of each tree
  importance = TRUE,   # Enable variable importance calculation
  na.action = na.omit
)
```


```{r}
# Model Overview
print(Randomforest_model0)

# Show the importance of the variable (numerical value)
importance(Randomforest_model0)

# Graphical display
varImpPlot(Randomforest_model0, main = "The importance of variables in stroke prediction")


# The marginal effect graph of age on stroke probability
partialPlot(Randomforest_model0, pred.data = balanced_train_data, x.var = "age", which.class = "1",
            main = "The marginal effect of age on stroke probability", xlab = "age", ylab = "probability of stroke")

```

```{r}
set.seed(123)

# 1. Classify the data
train_index <- createDataPartition(stroke_data$stroke, p = 0.7, list = FALSE)
train_data <- stroke_data[train_index, ]
test_data <- stroke_data[-train_index, ]

# Ensure that all categorical variables are factor
categorical_vars <- c("gender", "ever_married", "work_type", "Residence_type", "smoking_status", "stroke")

train_data[categorical_vars] <- lapply(train_data[categorical_vars], as.factor)
test_data[categorical_vars] <- lapply(test_data[categorical_vars], as.factor)  # Add this line

# 2. Balanced training set
balanced_train_data <- ROSE(stroke ~ ., data = train_data, seed = 1)$data

# 3. Force conversion to factors (ensure classification)
balanced_train_data$stroke <- factor(balanced_train_data$stroke, levels = c(0, 1))
train_data$stroke <- factor(train_data$stroke, levels = c(0, 1))
test_data$stroke <- factor(test_data$stroke, levels = c(0, 1))

# 4. training model
Randomforest_model0 <- randomForest(
  stroke ~ age + hypertension + heart_disease + ever_married + work_type +
    Residence_type + avg_glucose_level + bmi + smoking_status + gender,
  data = balanced_train_data,
  ntree = 500,
  mtry = 3,
  importance = TRUE,
  na.action = na.omit
)

# 5. forecast
pred_class <- predict(Randomforest_model0, newdata = test_data)
pred_prob <- predict(Randomforest_model0, newdata = test_data, type = "prob")[, "1"]

# 6. evaluate
conf_mat <- confusionMatrix(pred_class, test_data$stroke)
print(conf_mat)

roc_obj <- roc(test_data$stroke, pred_prob)
auc_val <- auc(roc_obj)
cat("AUC =", round(auc_val, 4), "\n")

plot(roc_obj, main = paste("ROC Curve - AUC =", round(auc_val, 4)))

# 7. Variable importance
varImpPlot(Randomforest_model0, main = "Variable Importance")

```
```{r}
# Obtain the predicted probability
pred_prob <- predict(Randomforest_model0, test_data, type = "prob")[, 2]

# Set a series of thresholds
thresholds <- seq(0.1, 0.9, by = 0.05)
f1_scores <- c()
recalls <- c()

# Calculate the F1 and Recall corresponding to each threshold in a loop
for (t in thresholds) {
  pred_class <- ifelse(pred_prob > t, 1, 0)
  f1 <- F1_Score(y_pred = pred_class, y_true = test_data$stroke, positive = "1")
  recall <- Recall(y_pred = pred_class, y_true = test_data$stroke, positive = "1")
  f1_scores <- c(f1_scores, f1)
  recalls <- c(recalls, recall)
}

# Visualize the changes of F1 and Recall with the threshold
plot(thresholds, f1_scores, type = "b", col = "blue", ylim = c(0, 1),
     ylab = "Score", xlab = "Threshold", main = "F1 & Recall vs Threshold")
lines(thresholds, recalls, type = "b", col = "red")
legend("topright", legend = c("F1 Score", "Recall"), col = c("blue", "red"), lty = 1)

# Find the optimal threshold (to maximize F1)
best_threshold <- thresholds[which.max(f1_scores)]
cat("Best threshold based on F1 Score:", best_threshold, "\n")

# Generate the final classification using the optimal threshold
final_pred <- ifelse(pred_prob > best_threshold, 1, 0)

# Output confusion matrix
confusionMatrix(as.factor(final_pred), as.factor(test_data$stroke), positive = "1")
```



#This model fixes other variables with a "benchmark sample" and only gradually changes the age to predict the probability of stroke at different ages.
```{r}

# 1
factor_vars <- c("gender", "hypertension", "heart_disease", "ever_married",
                 "work_type", "Residence_type", "smoking_status")
# 2
for (var in factor_vars) {
  stroke_data[[var]] <- factor(stroke_data[[var]])
}
train_index <- createDataPartition(stroke_data$stroke, p = 0.7, list = FALSE)
train_data <- stroke_data[train_index, ]
test_data <- stroke_data[-train_index, ]


 Randomforest_model <- randomForest(
   stroke ~ age + hypertension + heart_disease + ever_married + work_type +
     Residence_type + avg_glucose_level + bmi + smoking_status + gender,
   data = balanced_train_data,
   ntree = 500,
   mtry = 3,
   importance = TRUE,
   na.action = na.omit
 )

# 3. Select a benchmark sample row without any omissions
ref_row <- stroke_data[complete.cases(stroke_data), ][1, ]

# 4. Create an age sequence
age_seq <- seq(min(stroke_data$age), max(stroke_data$age), by = 1)

# 5. Construct a test dataset that only changes the age, while maintaining the baseline values for other variables
age_test_data <- do.call(rbind, lapply(age_seq, function(a) {
  temp <- ref_row
  temp$age <- a
  temp
}))

# 6. Delete useless columns (such as labels and redundant columns)
age_test_data$stroke <- NULL
age_test_data$age_group <- NULL 

# 7. Ensure that the test data factor variable levels are exactly the same as the training data
for (var in factor_vars) {
  age_test_data[[var]] <- factor(as.character(age_test_data[[var]]), levels = levels(stroke_data[[var]]))
}

# 8. Ensure that the types of numerical variables are correct
num_vars <- c("age", "avg_glucose_level", "bmi")
for (var in num_vars) {
  age_test_data[[var]] <- as.numeric(age_test_data[[var]])
}

# 9. Only retain the predictor variables used by the model (in the same order)
vars_in_model <- c("age", "hypertension", "heart_disease", "ever_married", "work_type",
                   "Residence_type", "avg_glucose_level", "bmi", "smoking_status", "gender")
age_test_data <- age_test_data[, vars_in_model]

# 10. The probability of stroke occurring is 1.
pred_probs <- predict(Randomforest_model, newdata = age_test_data, type = "prob")[, "1"]

# 11. Organize the prediction results into a data frame
age_effect <- data.frame(age = age_seq, stroke_prob = pred_probs)

# Smooth the "stroke_prob" value, with a window size of 5
age_effect$smoothed_prob <- rollmean(age_effect$stroke_prob, k = 5, fill = NA)

# Calculate the first-order difference
age_effect$diff_prob <- c(NA, diff(age_effect$smoothed_prob))

# Set a threshold for "significant increase", such as greater than 0.015
threshold <- 0.015

# Find the first position that exceeds the threshold
rise_start_index <- which(age_effect$diff_prob > threshold)[1]
rise_start_age <- age_effect$age[rise_start_index]

# Find the peak of the stroke probability
peak_index <- which.max(age_effect$stroke_prob)
peak_age <- age_effect$age[peak_index]
peak_prob <- age_effect$stroke_prob[peak_index]

# The graph shows the relationship between age and stroke probability
# ggplot(age_effect, aes(x = age, y = stroke_prob)) +
# geom_line(color = "blue") +
# labs(title = "The influence of age on the stroke probability",
# x = "age", y = "Predict the probability of stroke occurrence") +
# theme_minimal()


## Find the starting point where the stroke probability increases significantly (use first-order difference to detect the biggest growth point) 
# diff_probs <- diff(age_effect$stroke_prob)
# rise_start_index <- which.max(diff_probs)
# rise_start_age <- age_effect$age[rise_start_index]
# 
# # Find the peak age of the stroke probability
# peak_index <- which.max(age_effect$stroke_prob)
# peak_age <- age_effect$age[peak_index]
# peak_prob <- age_effect$stroke_prob[peak_index]
# 
# # Draw
# ggplot(age_effect, aes(x = age, y = stroke_prob)) +
#   geom_line(color = "blue") +
#   geom_vline(xintercept = rise_start_age, linetype = "dashed", color = "red") +
#   geom_vline(xintercept = peak_age, linetype = "dashed", color = "darkgreen") +
#   annotate("text", x = rise_start_age, y = 0.25,
#            label = paste0("Starting from", rise_start_age, " years old \nthe probability of stroke increases significantly."),
#            color = "red", angle = 90, hjust = -0.1, size = 4.5) +
#   annotate("text", x = peak_age, y = peak_prob + 0.05,
#            label = paste0("From", peak_age, " years old\nThe probability of a stroke reaches its peak"),
#            color = "darkgreen", angle = 90, hjust = -0.1, size = 4.5) +
#   labs(title = "The influence of age on the stroke probability",
#        x = "age", y = "Predict the probability of stroke occurrence") +
#   theme_minimal()

# reloading
ggplot(age_effect, aes(x = age, y = stroke_prob)) +
  geom_line(color = "blue") +
  geom_vline(xintercept = rise_start_age, linetype = "dashed", color = "red") +
  geom_vline(xintercept = peak_age, linetype = "dashed", color = "darkgreen") +
  annotate("text", x = rise_start_age, y = 0.7,
           label = paste0(rise_start_age, " years old"),
           color = "red", angle = 0, hjust = 0.5, size = 4.5) +
  annotate("text", x = peak_age, y = min(peak_prob + 0.02, 0.95),
           label = paste0(peak_age, " years old"),
           color = "darkgreen", angle = 0, hjust = 0.5, size = 4.5) +
  labs(title = "The influence of age on the probability of stroke",
       x = "age", y = "probability of stroke") +
  theme_minimal()


```

```{r}
# Obtain the predicted probability (the probability of stroke=1)
pred_prob <- predict(Randomforest_model, test_data, type = "prob")[, 2]

# Set a series of thresholds
thresholds <- seq(0.1, 0.9, by = 0.05)
f1_scores <- c()
recalls <- c()

# Calculate the F1 and Recall corresponding to each threshold in a loop
for (t in thresholds) {
  pred_class <- ifelse(pred_prob > t, 1, 0)
  f1 <- F1_Score(y_pred = pred_class, y_true = test_data$stroke, positive = "1")
  recall <- Recall(y_pred = pred_class, y_true = test_data$stroke, positive = "1")
  f1_scores <- c(f1_scores, f1)
  recalls <- c(recalls, recall)
}

# Visualize the changes of F1 and Recall with the threshold
plot(thresholds, f1_scores, type = "b", col = "blue", ylim = c(0, 1),
     ylab = "Score", xlab = "Threshold", main = "F1 & Recall vs Threshold")
lines(thresholds, recalls, type = "b", col = "red")
legend("topright", legend = c("F1 Score", "Recall"), col = c("blue", "red"), lty = 1)

# Find the optimal threshold (to maximize F1)
best_threshold <- thresholds[which.max(f1_scores)]
cat("Best threshold based on F1 Score:", best_threshold, "\n")

# Generate the final classification using the optimal threshold
final_pred <- ifelse(pred_prob > best_threshold, 1, 0)

# Output confusion matrix
confusionMatrix(as.factor(final_pred), as.factor(test_data$stroke), positive = "1")
```


```{r}
# Training data structure
str(balanced_train_data)
# Predicted data structure
str(age_test_data)

```




```{r}
# Calculate the first-order difference
diff_probs <- diff(pred_probs)

# Find the biggest growth point
max_diff_index <- which.max(diff_probs)
age_critical_point <- age_seq[max_diff_index]

cat("The age critical point at which the risk surges is approximately:", age_critical_point, "\n")

df_diff <- data.frame(
  age = age_seq[-1],  # Because the length decreases by 1 after diff
  diff_prob = diff_probs
)

ggplot(df_diff, aes(x = age, y = diff_prob)) +
  geom_line(color = "red", size = 1) +
  geom_vline(xintercept = age_critical_point, linetype = "dashed", color = "blue") +
  labs(title = "Age risk jump rate", x = "age", y = "Growth rate of stroke probability") +
  theme_minimal()

```

#The interaction curve of age × smoking status
```{r}
# Variable value
age_seq <- seq(min(stroke_data$age), max(stroke_data$age), by = 1)
smoking_levels <- levels(stroke_data$smoking_status)

#Construct the test data set
pred_list <- lapply(smoking_levels, function(smoke_status) {
  # Select a typical individual (benchmark row)
  ref_row <- stroke_data[1, ]
  ref_row$smoking_status <- smoke_status  # Fixed smoking status
  
  # Construct data according to age changes
  temp_data <- do.call("rbind", lapply(age_seq, function(a) {
    temp <- ref_row
    temp$age <- a
    temp
  }))
  
  # Variable type conversion (keeping consistent)
  factor_vars <- c("hypertension", "heart_disease", "ever_married",
                   "work_type", "Residence_type", "smoking_status", "gender")
  for (var in factor_vars) {
    temp_data[[var]] <- factor(temp_data[[var]], levels = levels(stroke_data[[var]]))
  }
  temp_data$age <- as.numeric(temp_data$age)
  temp_data$avg_glucose_level <- as.numeric(temp_data$avg_glucose_level)
  temp_data$bmi <- as.numeric(temp_data$bmi)
  
  temp_data$stroke <- NULL
  
  # prediction probability
  pred_prob <- predict(Randomforest_model, newdata = temp_data, type = "prob")[, "1"]
  
  data.frame(age = age_seq, smoking_status = smoke_status, stroke_prob = pred_prob)
})

# combined result
df_pred <- do.call(rbind, pred_list)

# graw
# ggplot(df_pred, aes(x = age, y = stroke_prob, color = smoking_status)) +
#   geom_line(size = 1.2) +
#   labs(title = "The interaction effect of age and smoking status on the stroke probability",
#        x = age, y = "Predict the probability of stroke occurrence", color = "Smoking status") +
#   theme_minimal()

# Add key inflection point calculation 
critical_points <- df_pred %>%
  group_by(smoking_status) %>%
  arrange(age) %>%
  mutate(diff_prob = c(NA, diff(stroke_prob))) %>%
  filter(!is.na(diff_prob)) %>%
  slice_max(diff_prob, n = 1)

# Print
print(critical_points[, c("smoking_status", "age", "diff_prob")], row.names = FALSE)


# Filter out "Unknown" and reset factor levels
df_pred_filtered <- df_pred %>% 
  filter(smoking_status != "Unknown") %>%
  mutate(smoking_status = factor(smoking_status, levels = c("formerly smoked", "never smoked", "smokes")))

critical_points_filtered <- critical_points %>%
  filter(smoking_status != "Unknown") %>%
  mutate(smoking_status = factor(smoking_status, levels = c("formerly smoked", "never smoked", "smokes")))

# Plot
ggplot(df_pred_filtered, aes(x = age, y = stroke_prob, color = smoking_status)) +
  geom_line(size = 1.2) +
  geom_vline(
    data = critical_points_filtered,
    aes(xintercept = age, color = smoking_status),
    linetype = "dashed", alpha = 0.6
  ) +
  geom_point(
    data = critical_points_filtered,
    aes(x = age, y = stroke_prob),
    color = "black", size = 3, shape = 1
  ) +
  labs(
    title = "Effect of Age and Smoking Status on Stroke Probability",
    x = "Age", y = "Stroke Probability",
    color = "Smoking Status",
    caption = paste("Critical ages:", paste(round(critical_points_filtered$age, 1), collapse = ", "), "years old")
  ) +
  theme_minimal()

```


```{r}
train_index <- createDataPartition(stroke_data$stroke, p = 0.7, list = FALSE)
train_data <- stroke_data[train_index, ]
test_data <- stroke_data[-train_index, ]
#step 1: Train the randomforest model (using the data balanced by ROSE)
set.seed(123)
RF_age_glucose <- randomForest(
  stroke ~ age + hypertension + heart_disease + ever_married + work_type +
    Residence_type + avg_glucose_level + bmi + smoking_status + gender,
  data = balanced_train_data,
  ntree = 500,
  mtry = 3,
  importance = TRUE
)

# step 2: Construct the variable combination grid of age × avg_glucose_level
age_seq <- seq(20, 85, by = 1)
glucose_seq <- seq(50, 250, by = 5)

# Create a variable combination grid
grid <- expand.grid(age = age_seq, avg_glucose_level = glucose_seq)

# Obtain a representative individual (ref_row) : used to fix other variables
ref_row <- stroke_data[1, ]
ref_row$age <- NULL
ref_row$avg_glucose_level <- NULL
ref_row$stroke <- NULL

# Build complete prediction data
predict_data <- cbind(grid, ref_row[rep(1, nrow(grid)), ])

# Ensure that the variable types are consistent with the training set
factor_vars <- c("hypertension", "heart_disease", "ever_married",
                 "work_type", "Residence_type", "smoking_status", "gender")
for (var in factor_vars) {
  predict_data[[var]] <- factor(predict_data[[var]], levels = levels(stroke_data[[var]]))
}
predict_data$bmi <- as.numeric(predict_data$bmi)

# step 3: Predict the probability of stroke occurrence
predict_data$stroke_prob <- predict(RF_age_glucose, newdata = predict_data, type = "prob")[, "1"]


```



```{r}
# 1. Select several points with the highest stroke_prob from the predicted data as key points
critical_points <- predict_data %>%
  slice_max(order_by = stroke_prob, n = 8)

# 2. Filter out the points where either "age" or "avg_glucose_level" is missing or exceeds the plotting coordinate range.
critical_points <- critical_points %>%
  filter(!is.na(age), !is.na(avg_glucose_level),
         age >= 20, age <= 85,
         avg_glucose_level >= 50, avg_glucose_level <= 250)

# 3. Redraw the graphics
ggplot(predict_data, aes(x = age, y = avg_glucose_level, z = stroke_prob)) +
  geom_contour_filled(breaks = seq(0, 1, by = 0.1), alpha = 0.8) +
  geom_contour(color = "white", alpha = 0.5, linewidth = 0.3) +
  scale_fill_viridis_d(option = "magma", direction = -1) +
  geom_point(data = critical_points, aes(x = age, y = avg_glucose_level), 
             shape = 21, fill = "yellow", color = "black", size = 3) +
  geom_text_repel(data = critical_points, 
                  aes(label = paste0(round(age), "y\n", round(stroke_prob * 100, 1), "%")),
                  size = 3, color = "white") +
  labs(
    title = "Stroke Probability Landscape by Age and Glucose Level",
    subtitle = "Contour bands show probability intervals",
    x = "Age (years)", 
    y = "Average Glucose Level (mg/dL)",
    fill = "Stroke Risk"
  ) +
  coord_cartesian(expand = TRUE) +
  theme_minimal() +
  theme(legend.position = "right")

```


```{r}
# step 1: Obtain the predicted probability
pred_prob <- predict(RF_age_glucose, test_data, type = "prob")[, 2]

# step 2: Set multiple thresholds and calculate F1 and Recall
thresholds <- seq(0.1, 0.9, by = 0.05)
f1_scores <- c()
recalls <- c()

for (t in thresholds) {
  pred_class <- ifelse(pred_prob > t, 1, 0)
  f1 <- F1_Score(y_pred = pred_class, y_true = test_data$stroke, positive = "1")
  recall <- Recall(y_pred = pred_class, y_true = test_data$stroke, positive = "1")
  f1_scores <- c(f1_scores, f1)
  recalls <- c(recalls, recall)
}

# step 3: Find the optimal threshold (maximum F1 score)
best_threshold <- thresholds[which.max(f1_scores)]
cat("Best threshold based on F1 Score:", best_threshold, "\n")

# step 4: Classification prediction is carried out using the optimal threshold
final_pred <- ifelse(pred_prob > best_threshold, 1, 0)

# step 5: Confusion matrix + Evaluation index
cm <- confusionMatrix(as.factor(final_pred), as.factor(test_data$stroke), positive = "1")
print(cm)

# step 6: Draw the ROC curve and calculate the AUC
roc_obj <- roc(test_data$stroke, pred_prob)
auc_value <- auc(roc_obj)
cat("AUC:", auc_value, "\n")

# Visualization of ROC curve
plot(roc_obj, col = "blue", main = "ROC Curve", print.auc = TRUE)

# step 7: Visualize F1 and Recall at different thresholds
plot_df <- data.frame(threshold = thresholds, F1 = f1_scores, Recall = recalls)

ggplot(plot_df, aes(x = threshold)) +
  geom_line(aes(y = F1, color = "F1 Score")) +
  geom_line(aes(y = Recall, color = "Recall")) +
  geom_vline(xintercept = best_threshold, linetype = "dashed", color = "black") +
  labs(title = "F1 Score & Recall vs. Threshold",
       x = "Threshold", y = "Metric Value",
       color = "Metric") +
  theme_minimal()


```
Confusion Matrix and Statistics

          Reference
Prediction   0   1
         0 1392  43
         1   27  31

Accuracy        : 0.9531
Kappa           : 0.4455
Sensitivity     : 0.4189
Specificity     : 0.9810
Precision       : 0.5345
F1 Score        : 0.4697
Balanced Acc    : 0.7000
AUC             : 0.9176





```{r}
#Used for age × a certain categorical variable
plot_age_interaction_critical_point <- function(data, rf_model, group_var, baseline_row_index = 1) {
  
  # Variable type convention
  factor_vars <- c("stroke", "hypertension", "heart_disease", "ever_married",
                   "work_type", "Residence_type", "smoking_status", "gender")
  
  # age range
  age_seq <- seq(min(data$age), max(data$age), by = 1)
  
  # Group list
  group_levels <- levels(data[[group_var]])
  
  # Construct the test data set, group and change age respectively, and fix group_var
  pred_list <- lapply(group_levels, function(g_level) {
    ref_row <- data[baseline_row_index, ]
    ref_row[[group_var]] <- g_level
    
    temp_data <- do.call("rbind", lapply(age_seq, function(a) {
      temp <- ref_row
      temp$age <- a
      temp
    }))
    
    # Type conversions remain consistent
    for (var in factor_vars) {
      if (var %in% names(temp_data)) {
        temp_data[[var]] <- factor(temp_data[[var]], levels = levels(data[[var]]))
      }
    }
    temp_data$age <- as.numeric(temp_data$age)
    if ("avg_glucose_level" %in% names(temp_data)) temp_data$avg_glucose_level <- as.numeric(temp_data$avg_glucose_level)
    if ("bmi" %in% names(temp_data)) temp_data$bmi <- as.numeric(temp_data$bmi)
    
    # Delete the target variable
    if ("stroke" %in% names(temp_data)) temp_data$stroke <- NULL
    
    # prediction probability
    pred_prob <- predict(rf_model, newdata = temp_data, type = "prob")[, "1"]
    
    data.frame(age = age_seq, 
               group = g_level, 
               stroke_prob = pred_prob,
               stringsAsFactors = FALSE)
  })
  
  df_pred <- do.call(rbind, pred_list)
  
  # Calculate the growth rate and inflection point
  critical_points <- df_pred %>%
    group_by(group) %>%
    arrange(age) %>%
    mutate(diff_prob = c(NA, diff(stroke_prob))) %>%
    filter(!is.na(diff_prob)) %>%
    slice_max(diff_prob, n = 1) %>%
    select(group, age, diff_prob)
  
  print(critical_points)
  
  # draw
  p <- ggplot(df_pred %>%
                group_by(group) %>%
                arrange(age) %>%
                mutate(diff_prob = c(NA, diff(stroke_prob))) %>%
                filter(!is.na(diff_prob)),
              aes(x = age, y = diff_prob, color = group)) +
    geom_line(size = 1) +
    geom_vline(data = critical_points, aes(xintercept = age, color = group),
               linetype = "dashed") +
    geom_point(data = critical_points, aes(x = age, y = diff_prob, color = group),
               size = 3) +
    labs(title = paste("age ×", group_var, "stroke probability growth rate and inflection point"),
         x = "age", y = "Stroke probability growth rate", color = group_var) +
    theme_minimal()
  
  print(p)
}

```

```{r}
plot_age_interaction_critical_point(stroke_data, Randomforest_model, "hypertension")
```

```{r}
plot_age_interaction_critical_point(stroke_data, Randomforest_model, "gender")
```

```{r}
plot_age_interaction_critical_point(stroke_data, Randomforest_model, "heart_disease")
```

```{r}
#plot_age_interaction_critical_point(stroke_data, Randomforest_model, "")
```



#Evaluate the performance of the randomforest model
```{r}

# Obtain the predicted probability (the probability of stroke=1)
pred_prob <- predict(Randomforest_model, test_data, type = "prob")[, 2]

# Set a series of thresholds
thresholds <- seq(0.1, 0.9, by = 0.05)
f1_scores <- c()
recalls <- c()

# Calculate the F1 and Recall corresponding to each threshold in a loop
for (t in thresholds) {
  pred_class <- ifelse(pred_prob > t, 1, 0)
  f1 <- F1_Score(y_pred = pred_class, y_true = test_data$stroke, positive = "1")
  recall <- Recall(y_pred = pred_class, y_true = test_data$stroke, positive = "1")
  f1_scores <- c(f1_scores, f1)
  recalls <- c(recalls, recall)
}

# Visualize the changes of F1 and Recall with the threshold
plot(thresholds, f1_scores, type = "b", col = "blue", ylim = c(0, 1),
     ylab = "Score", xlab = "Threshold", main = "F1 & Recall vs Threshold")
lines(thresholds, recalls, type = "b", col = "red")
legend("topright", legend = c("F1 Score", "Recall"), col = c("blue", "red"), lty = 1)

# Find the optimal threshold (to maximize F1)
best_threshold <- thresholds[which.max(f1_scores)]
cat("Best threshold based on F1 Score:", best_threshold, "\n")

# Generate the final classification using the optimal threshold
final_pred <- ifelse(pred_prob > best_threshold, 1, 0)

# Output confusion matrix
confusionMatrix(as.factor(final_pred), as.factor(test_data$stroke), positive = "1")

```
